---
title: Female ASA Presidents
author: JVA
date: '2016-04-22'
slug: female-asa-presidents
categories:
  - R
tags: [rolling]
---

I recently received an e-mail from the American Statistical Association ([ASA](https://www.amstat.org)) about the [Helen Walker Society](https://www.amstat.org/giving/helenwalkersociety.cfm).
The e-mail listed the 14 *"distinguished women leaders who have held the office of ASA president"* going back to Helen Walker in 1944.  

Really?  Fourteen? All the way back to 1944?  Wow.

I thought it would be interesting to visualize the percentage of female ASA presidents over time.  I found a listing of ASA presidents on [Wikipedia]("https://en.wikipedia.org/wiki/List_of_presidents_of_the_American_Statistical_Association") and used the data to create a timeline.

```{r, message=FALSE, warning=FALSE}
library(rvest)
library(plyr)
library(DT)
library(zoo)
library(tidyverse)

pres1 <- html("https://en.wikipedia.org/wiki/List_of_presidents_of_the_American_Statistical_Association") %>% 
  html_node("#mw-content-text") %>%
  html_text() %>%
  strsplit("\n")
pres2 <- pres1[[1]]
pres3 <- pres2[grepl("^\\d{4}", unlist(pres2))]

df <- data.frame(
  Year = as.numeric(substring(pres3, 1, 4)),
  President = trimws(gsub("\\d|-", "", pres3))
  ) %>%
  mutate(
    Number = seq_along(President),
    Sex = ifelse(Year %in% c(1944, 1952, 1956, 1980, 1987, 1989, 1992, 1996, 
      2006, 2007, 2009, 2011, 2013, 2016, 2018), "Female", "Male")
  ) %>%
  select(Number, Year, President, Sex)

datatable(df, rownames=FALSE,
  caption="Presidents of the American Statistical Association, listed by their first year of service")
```

I then calculated the rolling 10-year percentage of ASA presidents that were female and plotted the results.

```{r, message=FALSE, warning=FALSE}
df2 <- left_join(data.frame(Year=min(df$Year):max(df$Year)), df, by="Year")
df3 <- colwise(na.locf, fromLast=TRUE)(df2) %>%
  mutate(
    mean10 = 100*rollmean(Sex=="Female", 10, align="right", fill=NA)
  )

ggplot(df3, aes(Year, mean10)) +
  geom_line(size=1) +
  ggtitle("Female ASA Presidents") +
  labs(y = "Female presidents  (% of past 10 y)") +
  theme_bw(base_size=16)
```


There were three female ASA presidents between 1944 and 1956, followed by a 23-year period during which all the ASA presidents were male (1957-1979).  Since then, the percentage of female ASA presidents has risen, reaching 50% for the first time during the 10-year period from 2004 to 2013.  Excellent.
