---
title: Checking Citations
author: 'JVA'
date: '2017-05-25'
slug: checking-citations
categories:
  - R
tags:
  - peer review
---

If you conduct peer reviews of manuscripts for publication, you will do the author(s) a great favor by double checking their citations. 

* Is every work cited in the body of their text listed in the references section?
* Is every work listed in the references section cited in the body of the text?

Checking for one-to-one correspondence manually is time consuming.  That's why it's so common to find mismatches.

You can make the job a bit easier using the `xCitation()` function of the [**jvamisc**](https://github.com/JVAdams/jvamisc) package.

```{r}
library(jvamisc)
xCitations
```

I based the function on code posted by Kay Cichini on 26 Mar 2012 on [theBioBucket](http://thebiobucket.blogspot.com/2012/03/how-to-extract-citation-from-body-of.html).  I wanted something simple to serve my purposes, so I built the function to take either a text file path or object as input, and the function returns a sorted list of the unique citations.  This makes it easy for me to (yes, this part is still manual) compare them to the works listed in the references.

```{r}
x <- paste("Yarmouth (1977) said something.",
 "Evidence of x (Barber and Jones 1991),",
 "y (House et al. 1982),",
 "and z (Smith 1990; Folger and Penn 2000).")
xCitations(txt=x)
```

Note that this function does __not__ capture citations in which the author is mentioned directly in the text and only the year is in parentheses, _Yarmouth (1977)_ in the example.
