---
title: Who Tweets about the Great Lakes?
author: JVA
date: '2017-08-26'
slug: who-are-the-major-great-lakes-tweeters
categories:
  - R
tags:
  - Twitter
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

Twitter is one way that I keep up to date on Great Lakes news.  Might I be missing out on some good Great Lakes tweeters?  I used the [R](https://www.r-project.org/about.html) package [rtweet](https://cran.r-project.org/package=rtweet) to investigate (see this [vignette](https://cran.r-project.org/web/packages/rtweet/vignettes/intro.html)).

```{r}
library(rtweet)
library(tidyverse)
library(RColorBrewer)
library(lubridate)

# look at the last ntwee
ntweets <- 18000
gltweets <- search_tweets('
  "#GreatLakes" OR "Great Lakes" OR
  "#LakeSuperior" OR "Lake Superior" OR
  "#LakeMichigan" OR "Lake Michigan" OR
  "#LakeHuron" OR "Lake Huron" OR
  "#LakeErie" OR "Lake Erie" OR
  "#LakeOntario" OR "Lake Ontario"', 
  n=ntweets, include_rts=FALSE, verbose=FALSE) 
goback <- signif(diff(range(gltweets$created_at)), digits=2)
```

First, I looked at the last `r format(ntweets, big.mark=",")` tweets that mentioned the Great Lakes.  These tweets go back `r signif(goback, 2)` `r units(goback)`.

Then I counted up the number of Great Lakes tweets for each tweeter.

```{r}
gltweeters <- gltweets %>%
  users_data() %>%
  count(screen_name, name, description, followers_count) %>%
  arrange(-n, -followers_count) %>%
  select(No.GL.tweets=n, Handle=screen_name, Name=name,
    Description=description, No.followers=followers_count)

library(DT)
datatable(gltweeters, rownames=FALSE)
```

Not as enlightening as I had hoped.  I wanted to standardize the number of Great Lakes tweets by the total number of tweets from each person over the same time period, but I couldn't figure out how to do that.
