---
title: Government Shutdowns
author: JVA
date: '2018-01-21'
slug: government-shutdowns
categories:
  - R
tags:
  - politics
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

Facing another U.S. government shutdown, I came across an article on  [Vox.com](https://www.vox.com/policy-and-politics/2018/1/19/16905584/government-shutdown-history-clinton-obama-explained) describing the 18 previous shutdowns.  The article stated that *"... they’re a relatively recent development ... the result of the Congressional Budget Act of 1974."*

I was curious to visualize the timing and the duration of the shutdowns.  Happily, the beginning and ending dates were reported in the article.  I hand entered the data (it seemed faster than scraping).  In terms of writing an **R** script, the only trick was creating date variables (`plot.From` and `plot.To`) for plotting that used a single fiscal year (FY2001).

```{r}
library(tidyverse)
library(scales)

df <- data.frame(
  Shut = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18),
  mm1 = c(9, 9, 10, 11, 9, 9, 11, 9, 12, 11, 9, 10, 10, 12, 10, 11, 12, 10),
  dd1 = c(30, 30, 31, 30, 30, 30, 20, 30, 17, 10, 30, 3, 16, 18, 5, 13, 5, 1),
  yyyy1 = c(1976, 1977, 1977, 1977, 1978, 1979, 1981, 1982, 1982, 1983, 1984,
    1984, 1986, 1987, 1990, 1995, 1995, 2013),
  mm2 = c(10, 10, 11, 12, 10, 10, 11, 10, 12, 11, 10, 10, 10, 12, 10, 11, 1,
    10),
  dd2 = c(11, 13, 9, 9, 18, 12, 23, 2, 21, 14, 3, 5, 18, 20, 9, 19, 6, 17),
  yyyy2 = c(1976, 1977, 1977, 1977, 1978, 1979, 1981, 1982, 1982, 1983, 1984,
    1984, 1986, 1987, 1990, 1995, 1996, 2013)
  ) %>%
  mutate(
    From = as.Date(paste(yyyy1, mm1, dd1, sep="-")),
    To = as.Date(paste(yyyy2, mm2, dd2, sep="-")),
    Duration = as.numeric(To - From),
    FY = ifelse(mm2 > 8, yyyy2+1, yyyy2),
    # put From and To on a single calendar year for plotting
    Plot.year1 = ifelse(mm1 > 8, 2000, 2001),
    Plot.year2 = ifelse(mm2 > 8, 2000, 2001),
    plot.From = as.Date(paste(Plot.year1, mm1, dd1, sep="-")),
    plot.To = as.Date(paste(Plot.year2, mm2, dd2, sep="-"))
  )

ggplot(df, aes(x=plot.From, xend=plot.To, y=FY, yend=FY)) +
  geom_segment(color="#85bb65", size=2) +
  geom_point() +
  geom_point(aes(x=plot.To)) +
  scale_x_date(labels=date_format("%b %d")) +
  labs(x="Date", y="Fiscal year",
    title="Timing of Past US Gov't Shutdowns") +
  theme_bw(base_size=14)
```

There have been 18 shutdowns so far (not including this latest one), but sometimes there was more than one shutdown in a year.  There have been shutdowns in 13 fiscal years since 1974; that's an average of 1 shutdown every `r signif((2018-1974)/13, 2)` years.  The average total duration of the shutdowns in a given fiscal year was `r signif(mean(df$Duration), 2)` days.

```{r}
df %>%
  group_by(FY) %>%
  summarise(ndays = sum(Duration)) %>%
  ggplot(aes(FY, ndays)) +
    geom_bar(stat="identity", fill="#85bb65", color="black") +
    labs(x="Fiscal year", y="Days",
      title="Duration of Past US Gov't Shutdowns") +
    theme_bw(base_size=14)
```
