---
title: Finishing the Race
author: JVA
date: '2018-04-21'
slug: finishing-the-race
categories:
  - R
tags:
  - sports
  - gender
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

[Why Men Quit and Women Don’t](https://www.nytimes.com/2018/04/20/opinion/boston-marathon-women-nurse.html).  The tweet caught my eye.

`r htmltools::HTML("{{< tweet 987502049867812864 >}}")`

I wanted to see for myself the percentage of women and men that finished the Boston Marathon.  I found the 2018 data here: [http://registration.baa.org/2018/cf/Public/iframe_Statistics.htm](http://registration.baa.org/2018/cf/Public/iframe_Statistics.htm).  I replaced the year in the URL to get earlier years back to 2010.  Then I scraped the data using [R](https://www.r-project.org/) and the package [rvest](https://CRAN.R-project.org/package=rvest).

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(rvest)
library(binom)
library(DT)
```

```{r eval=FALSE}
library(tidyverse)
library(rvest)
library(binom)
library(DT)
```

```{r}
prefix <- "http://registration.baa.org/"
suffix <- "/cf/Public/iframe_Statistics.htm"

Year <- 2010:2018
# create empty list for results
results <- setNames(vector("list", length(Year)), Year)
for(i in 1:9) {
  # the 2018 data were in different rows than the other years
  getrows <- if(i<9) 4:5 else 6:7
  df <- read_html(paste0(prefix, Year[i], suffix)) %>%
    html_nodes("table") %>%
    # grab the 4th table
    .[4] %>%
    html_table(fill=TRUE) %>%
    # extract the first element from the list
    .[[1]] %>%
    # extract given rows and columns from the data frame
    .[getrows, c(1, 3:4)]
  # assign names (again, to deal with the different 2018 data)
  names(df) <- c("Gender", "Starters", "Finishers")
  results[[i]] <- df
}

# convert list to data frame
race <- bind_rows(results, .id="Year") %>%
  # convert numbered columns to numeric
  type_convert() %>%
  # convert Gender to a factor
  mutate(
    Gender = factor(Gender, level=c("female", "male"),
      label=c("Female", "Male"))
  )
```

I used the `binom.confint()` function from the [binom](https://CRAN.R-project.org/package=binom) package to calculate exact binomial confidence intervals around the percentage of starters that finished the race.  Then I plotted the percentage of finishers, by gender, over time.  (I did not include the data from 2013, when the race was cut short by the bombing.)

```{r}
prop <- with(race, binom.confint(Finishers, Starters, method="exact"))

raceprop <- bind_cols(race, 100*prop[, 4:6]) %>%
  rename(
    PctFinishers=mean,
    PctFinLo=lower,
    PctFinHi=upper
  ) %>%
  mutate(
    # set 2013 finishers to missing, marathon was cut short by bombing
    Finishers = ifelse(Year==2013, NA, Finishers),
    PctFinishers = ifelse(Year==2013, NA, PctFinishers),
    PctFinLo = ifelse(Year==2013, NA, PctFinLo),
    PctFinHi = ifelse(Year==2013, NA, PctFinHi)
  )

ggplot(raceprop, aes(Year, PctFinishers, color=Gender)) +
  geom_errorbar(aes(ymin=PctFinLo, ymax=PctFinHi), width=0.2, alpha=0.5,
    position=position_dodge(width=0.2)) +
  geom_point() +
  geom_line(size=1) +
  labs(y="Finishers  (% of Starters)") +
  theme_bw(base_size=14)
```

In most years, a higher percentage of males finish the race than females.  The two exceptions were **2012**, which was *"an unusually hot 86-degree day"* and **2018**, which had *"horizontal rain and freezing temperatures."*  Interesting.

```{r}
datatable(raceprop, options=list(pageLength=25)) %>%
  formatCurrency(c("Starters", "Finishers"), currency="", digits=0) %>%
  formatRound(c("PctFinishers", "PctFinLo", "PctFinHi"), 1)
```

