---
title: Chronic Wasting Disease in Wisconsin Deer
author: JVA
date: '2019-12-24'
slug: chronic-wasting-disease-in-wisconsin-deer
categories:
  - R
tags:
  - web scraping
  - map
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

Do you deer hunt in Wisconsin?  Have you had your deer tested for chronic wasting disease?  I recently learned the results of the test on my deer and was curious about the results statewide.

# Bringing in the data

The county data are available online at [dnr.wi.gov](https://dnr.wi.gov/wmcwd/Summary/County).  I used [SelectorGadget](https://cran.r-project.org/web/packages/rvest/vignettes/selectorgadget.html) to identify the columns of data I wanted from each page and the **rvest** package to do the extraction, looping through the pages with different years of data.

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(rvest)
library(maps)
library(DT)
```

```{r eval=FALSE}
library(tidyverse)
library(rvest)
library(maps)
library(DT)
```

```{r}
prefix <- "https://dnr.wi.gov/wmcwd/Summary/YearCounty/"
Years <- c("1999-2001", 2002:2019)

# create empty list for results
results <- setNames(vector("list", length(Years)), Years)

# loop through the years to fill up the list
for(i in 1:19) {
  County <- read_html(paste0(prefix, Years[i])) %>%
    html_nodes("td:nth-child(1)") %>%
    html_text(trim=TRUE)
  Analyzed <- read_html(paste0(prefix, Years[i])) %>%
    html_nodes("td:nth-child(3)") %>%
    html_text(trim=TRUE) %>%
    as.numeric()
  Positive <- read_html(paste0(prefix, Years[i])) %>%
    html_nodes("td:nth-child(4)") %>%
    html_text(trim=TRUE) %>%
    as.numeric()
  results[[i]] <- data.frame(County, Analyzed, Positive, 
    stringsAsFactors=FALSE)
}

# convert list to data frame
cwd <- bind_rows(results, .id="Years") %>%
  mutate(
    Year = as.numeric(ifelse(Years=="1999-2001", "2000", Years)),
    Pos.pct = 100*Positive/Analyzed,
    County = ifelse(County=="Saint Croix", "St Croix", County),
    county = casefold(County, upper=FALSE)
  ) %>%
  filter(County!="Unknown")
```

In preparation for mapping the results, I also downloaded county mapping data for Wisconsin.

```{r}
WI <- map_data("county") %>%
  filter(region=="wisconsin") %>%
  rename(county=subregion)

WIcenters <- WI %>%
  group_by(county) %>%
  summarise(long=mean(range(long)), lat=mean(range(lat)))
```

# Summarizing the data

Then I summarized the test data, determining the first year that CWD was detected in each county and the percentage of positive results over all tests during 1999-2019 and just during 2019. 

```{r}
# summarize 2019 data
cwd19 <- cwd %>%
  filter(Year==2019) %>%
  select(County, Analyzed.2019=Analyzed, Positive.2019=Positive, 
    Pos.pct.2019=Pos.pct)

# summarize all years of data
cwdall <- cwd %>%
  mutate(
    YP=ifelse(Positive>0, Year, NA)
  ) %>%
  group_by(County, county) %>%
  summarize(
    First.Detected=min(YP, na.rm=TRUE), 
    Analyzed.Total=sum(Analyzed), 
    Positive.Total=sum(Positive)) %>%
  mutate(
    Pos.pct.Total=100*Positive.Total/Analyzed.Total
  ) %>%
  mutate(
    First.Detected=ifelse(is.finite(First.Detected), First.Detected, NA)
  ) %>%
  full_join(cwd19) %>%
  full_join(WIcenters) %>%
  mutate(
    # mapit is missing for never detected and % positive for 2019 otherwise
    mapit = ifelse(Pos.pct.Total>0, Pos.pct.2019, NA)
  ) %>%
  arrange(-Pos.pct.Total, County)

cwdall %>%
  select(-county, -lat, -long, -mapit) %>%
  datatable() %>%
  formatCurrency(c("Analyzed.Total", "Positive.Total", 
    "Analyzed.2019", "Positive.2019"), currency="", digits=0) %>%
  formatRound(c("Pos.pct.Total", "Pos.pct.2019"), 1)
```

Since 1999, **`r sum(!is.na(cwdall$mapit))`** of 72 Wisconsin counties have tested positive for CWD, including **`r sum(cwdall$First.Detected==2019, na.rm=TRUE)`** counties that tested positive for the first time in 2019: **`r paste(cwdall$County[cwdall$First.Detected==2019 & !is.na(cwdall$First.Detected)], collapse=", ")`**.

# Mapping the data

I used the data to create a county map showing the percentage of positive results in 2019, the counties that have never tested positive for CWD, and the first year of detection.

```{r, fig.width=6, fig.height=6}
WIplus <- cwdall %>%
  select(-lat, -long) %>%
  full_join(WI)

ggplot(WIplus, aes(long, lat)) +
  # county polygons
  geom_polygon(aes(group=group, fill=mapit), colour="white") +
  # black text on all counties with any detections since 1999
  geom_text(data=cwdall, size=3.5, fontface="bold", 
    aes(long, lat, label=substring(First.Detected, 3, 4))) +
  # writeover white text on dark colored counties
  geom_text(data=filter(cwdall, Pos.pct.Total<1.1), size=3.5, fontface="bold", 
    color="white", aes(long, lat, label=substring(First.Detected, 3, 4))) +
  scale_fill_viridis_c(name="% Positive\nin 2019") +
  coord_map() +
  # remove default ggplot background, grid, and axes
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(title="Chronic Wasting Disease in Wisconsin Deer", 
    subtitle="Shading represents the percentage of deer that tested positive for CWD.\nCounties that have never tested positive since 1999 are in gray.\nLabels indicate the first year CWD was detected.")
```

# Plotting the data

Finally, to see the progression of the infection in the deer population over time, I plotted the percentage of positive results for every county that ever had a positive test.  I only show percentages for years with at least 8 deer tested in each county.  (*The combined data from 1999-2001 is plotted at 2000.*)

```{r}
cwd8 <- cwd %>%
  filter(Analyzed>=8 & County %in% cwdall$County[!is.na(cwdall$mapit)]) %>%
  mutate(
    County=factor(County, levels=cwdall$County)
  )

ggplot(cwd8, aes(Year, Pos.pct)) +
  geom_smooth(color="cyan", size=1, method="loess", span=2) +
  geom_point(alpha=0.5) +
  scale_x_continuous(breaks=seq(2000, 2019, 5), 
    labels=substring(seq(2000, 2019, 5), 3, 4)) +
  facet_wrap(~ County) +
  labs(y="% Positive for CWD") +
  theme_bw(base_size=12)
```

```{r, eval=FALSE, echo=FALSE}
# two other interesting plots that I didn't share

ggplot(cwdall, aes(First.Detected, Pos.pct.Total, size=Pos.pct.2019, color=Pos.pct.2019)) +
  geom_point()

ggplot(cwd, aes(Analyzed, Positive, color=Pos.pct)) +
  geom_point() + 
  scale_x_log10() +
  scale_y_log10()
```
